---
title: "Coronavirus Snapshot"
author: "Author: Grant Chalmers"
date: "Created 5 Mar, 2020; Last Update `r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    df_print: paged
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = "#>",
  fig.align = "centre",
  fig.height = 7,
  fig.width = 9,
  message = FALSE,
  warning = FALSE,
  error = FALSE)

knitr::knit_hooks$set(inline = function(x) {
prettyNum(x, big.mark=",")
})

options(tibble.print_max = 5, tibble.print_min = 5)
```

```{r load_libraries, include=FALSE}
library(rvest)
library(janitor)
library(tidyverse)
library(gghighlight)
library(httr)
library(jsonlite)
library(lubridate)
library(scales)

```



```{r setup_basic, include=FALSE, echo=FALSE}
# Setup data
# Remove scientific notation
options(scipen=999)

url <- "https://www.worldometers.info/coronavirus/"

# Set theme for ggplot
theme_set(theme_minimal())
# //*[@id="main_table_countries_today"]

```



```{r gather_clean, echo=FALSE, include=FALSE}
coronavirus <- url %>% 
  xml2::read_html() %>% 
  html_node(xpath = '//*[@id="main_table_countries_today"]') %>% 
  html_table()


# Use Janitor function to clean column names
coronavirus <-  coronavirus %>% clean_names()
# Review
coronavirus %>% glimpse()
# Clean up commas from HTML numeric format
coronavirus <- coronavirus %>% 
  mutate_all(funs(str_replace(., ",", "")))
# Convert to numeric
corona <- coronavirus %>% select(-country_other) %>% 
  mutate_if(is.character, as.numeric)

# Add countries back in
corona <- corona %>% add_column(coronavirus$country_other)
# Review
glimpse(corona)

# Fix name of country column and change order so it is first, calc total_deaths
# before converting to long (tidy) format
corona <- corona %>%
  select(country = `coronavirus$country_other`, everything()) %>% 
  mutate(total_deaths = total_cases - (active_cases + total_recovered),
         death_rate = round((total_deaths / total_cases * 100), 2))
# Review
glimpse(corona)

head(corona)

# Gather all data, tidy and standardise country names in prep for joining
all_corona <- corona %>% 
  filter(!country %in% c("Total:", "Diamond Princess")) %>% 
  pivot_longer(-country, names_to = "observation", values_to = "count") %>% 
  mutate(obs_ord = factor(observation,
                          levels=c("total_cases", "new_cases", "active_cases",
                                   "total_recovered", "total_deaths",
                                   "serious_critical", "death_rate",
                                   "tot_cases_1m_pop"),
                          labels=c("Total Cases", "New Cases", "Active Cases",
                                   "Total Recovered", "Total Deaths",
                                   "Serious Critical", "Death Rate",
                                   "Total Cases per 1m Population")),
         country = recode(country,
                          "S. Korea" = "South Korea",
                          "USA" = "United States",
                          "U.K." = "United Kingdom",
                          "U.A.E." = "United Arab Emirates",
                          "Czechia" = "Czech Republic",
                          "State of Palestine" = "Palestine"),
         tstamp = Sys.Date()) %>% 
  select(tstamp, country, observation, obs_ord, count)

# Get continent mapping from here: https://ourworldindata.org/world-region-map-definitions
# continents <- read_csv("~/R/coronavirus/data/continents-our-world-in-data.csv")
# 
# # Join by Country
# all_corona <- all_corona %>% left_join(continents, by = "country")

glimpse(all_corona)

```

```{r write_csv, echo=FALSE, include=FALSE}
# Write csv for time series analysis, later...
data_path <- "~/R/coronavirus/data/"   # path to the data
sys_date <- Sys.Date()
filename <- paste0(data_path,"corona", "_", sys_date, ".csv")

# Write to csv
# Write to csv
if (exists("all_corona")) {
  write_csv(all_corona, filename)
}
```


### Review overall totals.

Note: all totals exclude the Diamond Princess.

```{r totals, echo=FALSE}
# Review overall totals

all_corona %>% filter(!is.na(count), !obs_ord %in% c("Death Rate", "Total Cases per 1m Population", NA)) %>%
  group_by(obs_ord) %>%
  summarise(grp_tot = sum(count)) %>% 
  select(Observation = obs_ord,
         Total = grp_tot) %>% knitr::kable(format.args = list(decimal.mark = '.', big.mark = ","))

```

### Australian totals..

```{r oz_totals, echo=FALSE}
all_corona %>% filter(!is.na(count), !is.na(obs_ord), country == "Australia") %>%
  group_by(country, obs_ord) %>%
  summarise(grp_tot = sum(count)) %>% 
  select(Country = country,
         Observation = obs_ord,
         Total = grp_tot) %>% 
   knitr::kable(format.args = list(decimal.mark = '.', big.mark = ","))
```

### Death Rates

Top 10 death rates. Newly infected countries may initially have a high rate,
which should decrease over time.

```{r death_rates, echo=FALSE}
# Top 10 death rate
all_corona %>% filter(!is.na(count), obs_ord == "Death Rate") %>%
  top_n(10, wt = count) %>% 
  select(Country = country,
         "Death Rate (%)" = count) %>% 
  arrange(desc(`Death Rate (%)`)) %>% 
  knitr::kable(format.args = list(decimal.mark = '.', big.mark = ","))
```

### Total cases per 1 million population


```{r tot_cases_1m_pop, echo=FALSE}
# Top 10 cases per 1m of population
all_corona %>% filter(!is.na(count), obs_ord == "Total Cases per 1m Population") %>%
  top_n(10, wt = count) %>% 
  select(Country = country,
         "Total Cases per 1m Population" = count) %>% 
  arrange(desc(`Total Cases per 1m Population`)) %>% 
  knitr::kable(format.args = list(decimal.mark = '.', big.mark = ","))

```

View a lollipop chart of the top 50 countries. 
Note: x-axis is log scale.

### Total Cases.

```{r plot_lollipop, echo=FALSE}
all_corona %>%
  filter(observation %in% c("tot_cases_1m_pop"),
         count > 0) %>% 
  group_by(observation) %>%
  top_n(n = 50, wt = count) %>% 
  ggplot(aes(x=fct_reorder(country, count), y=count)) +
  geom_point(size=3, colour = "plum") + 
  geom_segment(aes(x=country, 
                   xend=country, 
                   y=0, 
                   yend=count), colour = "plum", size = 0.75) +
  # gghighlight(country == "Australia", unhighlighted_colour = "gray75") +
  coord_flip() +
  # scale_y_log10(labels = scales::comma) +
  # facet_wrap(~ obs_ord, scales = "free") +
  labs(title = "Coronavirus Total Cases per 1m Population (top 50)",
       y = "Count", x = NULL,
       caption = "Source: @GrantChalmers | https://www.worldometers.info/coronavirus/") +
  theme(plot.title = element_text(size = 11, face = "bold"),
        axis.text.x = element_text(size = 10, angle = 00),
        axis.title.x = element_text(size = 9),
        strip.text.x = element_text(size = 10, colour = "darkgreen", face = "bold"),
        axis.text.y = element_text(size = 8.5, colour = "darkblue"),
        axis.title = element_text(size = 11), legend.position = "none",
        plot.caption = element_text(size = 7.5, color = "gray50", face = "italic"),
        legend.title = element_blank())

# ggsave("lollipop_facet.png", path = "images")

```

Time series trend (log scale) for selected countries. 

*Highlight 9 random countries out of the top 100 for some variety each day!!*

```{r line_plot, echo=FALSE}
data_path <- "~/R/coronavirus/data/"   # path to the data

files <- dir(data_path, pattern = "*.csv") # get file names

coronadf <- files %>%
  map_df(~ read_csv(file.path(data_path, .),
                    col_types = cols(tstamp = col_date())))

top_100_new_cases <- coronadf %>% 
  filter(observation %in% c("new_cases"),
         count > 0) %>%
  group_by(country) %>%
  summarise(grp_total = sum(count)) %>% 
  top_n(n = 100, wt = grp_total) %>% pull(country) %>% unique()

coronadf %>%
  filter(observation	%in% c("tot_cases_1m_pop"),
         country %in% top_100_new_cases,
         !is.na(count)) %>% 
  ggplot(aes(x = tstamp, y = count)) +
  geom_line(aes(colour = country), size = 1) +
  gghighlight(country %in% sample(top_100_new_cases, 9),     # try randon sample
  # gghighlight(country %in% c("Australia", "Kazakhstan", "Canada", "Namibia", "Ukraine",
  #                            "Niger", "China", "Russia", "Uzbekistan", "United States"),
              unhighlighted_colour = ggplot2::alpha("grey", 0.7),
              use_group_by = FALSE, label_key = country, label_params = list(size = 3, alpha = 0.75)) +
  # scale_y_continuous(trans = 'log10') +
  scale_y_log10(labels = scales::comma) +
  scale_x_date(date_breaks = "3 day", date_labels = "%d %b") +
  labs(title = "Coronavirus Trend - Top Uranium Producers",
       y = "Total Cases per 1m Population (log scale)", x = "",
       caption = "Source: @GrantChalmers | https://www.worldometers.info/coronavirus/") + 
  theme_light() +
  theme(plot.title = element_text(size = 11, face = "bold"),
        axis.text.x = element_text(size = 10, angle = 00),
        axis.title.x = element_text(size = 9),
        strip.text.x = element_text(size = 10, colour = "darkgreen", face = "bold"),
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        axis.title = element_text(size = 11), legend.position = "bottom",
        plot.caption = element_text(size = 7.5, color = "gray50", face = "italic"),
        legend.title = element_blank())

ggsave("total_cases_per_1m_log.png", path = "images", width = 8, height = 6)

```

### New Cases.

```{r lollipop_new_cases, echo=FALSE}
all_corona %>%
  filter(observation %in% c("new_cases"),
         count > 0) %>%
  group_by(observation) %>%
  top_n(n = 50, wt = count) %>%
  ggplot(aes(x=fct_reorder(country, count), y=count)) +
  geom_point(size=3, colour = "plum") + 
  geom_segment(aes(x=country, 
                   xend=country, 
                   y=0, 
                   yend=count), colour = "plum", size = 0.75) +
  # gghighlight(country == "Australia", unhighlighted_colour = "gray75") +
  # if(any(all_corona$country == "Australia")) #gghighlight(country == "Australia", unhighlighted_colour = "gray75")
  coord_flip() +
  scale_y_log10(labels = scales::comma) +
  # facet_wrap(~ obs_ord, scales = "free") +
  labs(title = "Coronavirus New Cases by Country (top 50)",
       y = "Count (log scale)", x = NULL,
       caption = "Source: @GrantChalmers | https://www.worldometers.info/coronavirus/") +
  theme(plot.title = element_text(size = 11, face = "bold"),
        axis.text.x = element_text(size = 10, angle = 00),
        axis.title.x = element_text(size = 9),
        strip.text.x = element_text(size = 10, colour = "darkgreen", face = "bold"),
        axis.text.y = element_text(size = 8.5, colour = "darkblue"),
        axis.title = element_text(size = 11), legend.position = "none",
        plot.caption = element_text(size = 7.5, color = "gray50", face = "italic"),
        legend.title = element_blank())
```

### Active Cases

```{r lollipop_active_cases, echo=FALSE}
all_corona %>%
  filter(observation %in% c("active_cases"),
         count > 0) %>% 
  group_by(observation) %>%
  top_n(n = 50, wt = count) %>%
  ggplot(aes(x=fct_reorder(country, count), y=count)) +
  geom_point(size=3, colour = "peru") + 
  geom_segment(aes(x=country, 
                   xend=country, 
                   y=0, 
                   yend=count), colour = "peru", size = 0.75) +
  gghighlight(country == "Australia", unhighlighted_colour = "gray75") +
  coord_flip() +
  scale_y_log10(labels = scales::comma) +
  # facet_wrap(~ obs_ord, scales = "free") +
  labs(title = "Coronavirus Active Cases by Country (top 50)",
       y = "Count (log scale)", x = NULL,
       caption = "Source: @GrantChalmers | https://www.worldometers.info/coronavirus/") +
  theme(plot.title = element_text(size = 11, face = "bold"),
        axis.text.x = element_text(size = 10, angle = 00),
        axis.title.x = element_text(size = 9),
        strip.text.x = element_text(size = 10, colour = "darkgreen", face = "bold"),
        axis.text.y = element_text(size = 8.5, colour = "darkblue"),
        axis.title = element_text(size = 11), legend.position = "none",
        plot.caption = element_text(size = 7.5, color = "gray50", face = "italic"),
        legend.title = element_blank())
```

### Total Recovered

```{r lollipop_recovered_cases, echo=FALSE}
all_corona %>%
  filter(observation %in% c("total_recovered"),
         count > 0) %>%
  group_by(observation) %>%
  top_n(n = 50, wt = count) %>%
  ggplot(aes(x=fct_reorder(country, count), y=count)) +
  geom_point(size=3, colour = "peru") + 
  geom_segment(aes(x=country, 
                   xend=country, 
                   y=0, 
                   yend=count), colour = "peru", size = 0.75) +
  gghighlight(country == "Australia", unhighlighted_colour = "gray75") +
  coord_flip() +
  scale_y_log10(labels = scales::comma) +
  # facet_wrap(~ obs_ord, scales = "free") +
  labs(title = "Coronavirus Recovered Cases by Country (top 50)",
       y = "Count (log scale)", x = NULL,
       caption = "Source: @GrantChalmers | https://www.worldometers.info/coronavirus/") +
  theme(plot.title = element_text(size = 11, face = "bold"),
        axis.text.x = element_text(size = 10, angle = 00),
        axis.title.x = element_text(size = 9),
        strip.text.x = element_text(size = 10, colour = "darkgreen", face = "bold"),
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        axis.text.y = element_text(size = 8.5, colour = "darkblue"),
        axis.title = element_text(size = 11), legend.position = "none",
        plot.caption = element_text(size = 7.5, color = "gray50", face = "italic"),
        legend.title = element_blank())
```

### Australian States

```{r oz_states, echo=FALSE}

# Gather Australian geos info.
get_corona <- GET("https://api.infotorch.org/api/covid19/statlist/?geos=NSW,VIC,QLD,WA,SA,TAS,ACT,NT&stat=confirmed")
get_text <- content(get_corona, "text")
get_json <- fromJSON(get_text, flatten = TRUE)
coronadf <- as_tibble(get_json)

# Drop columns, unest data (list) column, spread, rename & export to csv
corona_au <- coronadf %>%
  select(-id, -geo_long) %>% 
  unnest(data) %>% 
  group_by(geo, x, y) %>% 
  select(state = geo,
         stat,
         date = x,
         value = y) %>%
  ungroup() %>% 
  mutate(date = as_date(date)) 

# corona_au %>% glimpse()

corona_au %>% 
  ggplot(aes(date, value, colour = state)) +
  geom_line(size = 1) +
  geom_text(data = corona_au %>% filter(date == max(date)),
            aes(x = date, y = value, label = state, group = state),
            position = position_dodge(0.9), size = 3, hjust = -0.25) +
  scale_y_continuous(labels = scales::comma) +
  scale_x_date(date_breaks = "3 day", date_labels = "%d %b") +
  scale_colour_brewer(palette = "Dark2") +
  labs(title = "Coronavirus Confirmed Cases Australian States",
       y = "Confirmed Cases", x = "", colour = "State",
       caption = "Source: @GrantChalmers | https://github.com/infotorch/covid19-australia-api") + 
  theme_light() +
  theme(plot.title = element_text(size = 11, face = "bold"),
        axis.text.x = element_text(size = 10, angle = 00),
        axis.title.x = element_text(size = 9),
        strip.text.x = element_text(size = 10, colour = "darkgreen", face = "bold"),
        plot.margin = margin(0.5, 0.5, 0.5, 0.5, "cm"),
        axis.title = element_text(size = 11), legend.position = "none",
        plot.caption = element_text(size = 7.5, color = "gray50", face = "italic"),
        legend.title = element_blank())

ggsave("corona_au_states.png", path = "images", width = 8, height = 6)


```



